\Large \bf Cyclone: Warp Speed Replication for Key Value Stores

### 2017-04-21 paper/paper.tex
Updates to paper.

% Paper number and no. of pages as author
\authorinfo{Paper \textbf{\#224}}{NN pages}

@@

In the absence of failures Figure~\ref{fig:ganged_ops} is straightforward. Once
replication is complete on all participating physical logs, the corresponding KV
store cores execute {\tt event\_replicated\_ganged\_op}. They use a bitmask for
the barrier setting the appropriate bit for themselves. Once all KV store cores
have arrived at the barrier the operation is executed. A distinguished leader
core (we use the minumum numbered participating KV store core) is
responsible for sending back the response to the client. Two types of failure
need to be considered. First, replication on the physical log corresponding to
one of the non-leader cores could fail. The
remaning cores would never progress past the barrier. To solve this the leader
core montors the remaning cores for their current view - which is the RAFT term
of the last executed log entry. If the view has moved past the view in the
packet, replication failed on that physical log. The leader then sets the bit
for the core that will no longer participate in the barrier to aid progress (a
technique we borrowed from lock-free data structures~\cite{lock-free}). It also
sets a failed flag to signal failure to the other participating cores. The
second failure case is that the leader core might not receive the operation if
it fails during replication from corresponding physical log. This case is
detected in a similar manner by the remaining cores by monitoring the published
view from the leader core in shared memory. We add that on a leadership change
due to timeout we enqueue and execute a nop on the physical log and all
associated KV store cores to ensure previous entries are committed - a corner
case in the RAFT protocol.

@@

\section{Evaluation}
\label{sec:evaluation}
We evaluate Cyclone on a 12 node cluster connected via a 10 GigE switch. Three
of the machines are equipped with 1.6TB Intel DC P3600 SSDs and 4*10 GigE
ports. The remaining nine machines do not have SSDs and have only one 10 GigE
port, serving as clients for most of the experiments. As with other
work~\cite{faast}, we use DRAM on the machines to proxy for NVDIMMs where
necessary - the persistent memory needed never exceeds 64 MB regardless of the
size of the key value store or second level log on flash. We divide the
evaluation into three parts. First, we evaluate Cyclone's performance with a
single level log as a pure software packet switch. Next, we evaluate performance
when adding a second level of log on flash. Finally, we evaluate performance
when integrated with Rocksdb as an alternative to Rocksdb's write ahead
log. Unless otherwise mentioned, we use a 60 byte header followed by a payload
for experiments. We log both the header and payload. In all cases the server
echoes the received entry (header and payload) back to the client. We fix the
number of application threads at 32 using at most 8 threads for running RAFT
instances.

Cyclone is built on top of DPDK to apply software packet switching techniques to
the log replication problem. We therefore begin by systematically evaluating
optimizations applied in Cyclone to replicate the top level log in
Figure~\ref{fig:network_opts} - with no payload. The y-axis reports latency seen
at the client (which means two network round trips with replication). Using
TCP/IP to replicate a RAFT log tops out at around 30K entries/s. Applying
optimizations improves performance by a large enough amount that we use a log
scale on the x-axis. Switching to DPDK (the line marked +DPDK) improves the
throughput by an order of magnitude to around 50K entries/s. Using batching
(line marked +batching) improves the performance further bringing us close to a
million entries/s. Scaling horizontally to 8 physical logs (+8 phy logs)
improves performance to close to 2M entries/s. Finally using all 4 ports on the
machine to replicate entries improves performance considerably to 6M
entries/s. In all performance improves by 200X over the TCP/IP single log
baseline. Cyclone also considerably improves the latency for replication, from
close to 100us with TCP/IP to around 30us at peak throughput.

@@

There are two factors that can have significant impact on Cyclone's
performance. First, the number of replicas dictates the outgoing message rate
from the leader replica and therefore increasing the replication factor can
decrease Cyclone's performance. Figure~\ref{fig:replicas} shows the impact of
varying replica count. Using only a single replica cuts out a network round trip
and shows the best unloaded latency (10 us) and peak throughput (near 10M
entries/s). Adding replicas decreases the peak throughput down to around 2M
entries/s. We note that a number of previous pieces of work~\cite{faast, farm}
use three replicas and therefore we focus on three replicas for the replicated
cases we consider below. The second factor that dictates Cyclone's performance
is the size of the log entry being replicated. Figure~\ref{fig:payload} shows
the effect of increasing the payload size from zero to 512 bytes. Peak
throughput drops from 6M entries/s to approximately 2M entries/s. At this
replication rate, the leader replica needs to transmit data at approximately 30
Gbit/s. Coupled with the cost of network headers all four 10 GigE ports are now
saturated and therefore Cyclone hits the network line rate bottleneck at this
point.

@@

We now turn our attention from the network component of Cyclone to the storage
one by adding the second level flashlog. We evaluate the impact of adding
storage to our hitherto pure packet switching scenario in
Figure~\ref{fig:flashlog}. Batching entries from the top level NVDIMM log to the
second level flashlog is clearly beneficial as adding flash storage at the
second level has almost no impact on peak performance in Cyclone for small
entries. The situation however changes for larger
entries. Figure~\ref{fig:flashlog_512} shows that using a 512 byte payload has a
significant impact on peak throughput - it drops to approximately 350K
ops/sec. This corresponds to around 50K 4KB IOPS to the SSD to write out the
flashlog pages. The peak for the drive is 160K IOPS using a queue depth
(concurrency) of 128. With 32 application threads we expect a lower peak
throughput around 40K IOPS explaining our bottleneck at 50K IOPS.  It is
possible to tune our observed performance further by aligning the flush boundary
to increase the number of outstanding requests - we do not do so in this paper,
keeping Cyclone agnostic to the exact size of entry being replicated. A final
point about Figure~\ref{fig:flashlog_512} is that once we are past the storage
bottleneck the latency spike is dramatic and large enough to trigger Cyclone's
failure detector and repeated retries from the clients. There are - therefore -
no points on the ``knee'' of the curve as in the pure packet switched one level
log case.

@@

\begin{figure}
\includegraphics[scale=0.6]{results2/flashlog_512.pdf}
\caption{Impact of adding second level log (512 bytes payload)}
\label{fig:flashlog_512}
\end{figure}

The final dimension we evaluate is of ganged operations. The primary purpose of
ganged operations is to avoid the need for distributed transactions to
manipulate what is a single shared memory image and therefore we were most
concerned about unloaded latency given the complexity of synchronizing different
replication quorums as well executing our rendezvous protocol on multiple
cores. We therefore setup an experiment where a single client reflecting the
unloaded case made ganged requests to the replicas. We varied the number of
cores participating in the request from one to the full complement of 32
application cores. Figure~\ref{fig:ganged} shows the results both using a single
level log as well as a two level log. The primary takeaway is that unloaded
latency increases slowly as we increase the number of active cores - to around
40 us from the baseline of 20 us. There are two causes of this. First, there is
a rapid increase from 1 to 8 active cores, the reason being that this
corresponds to an increase in the number of quorums that must be synchronized
on. There is some amount of synchronization involved in the userspace DPDK
driver for access to common resources for accessing the NIC and this affects the
ability to simultaneously issues replication messages on all quorums. The
smaller contribution to increasing latency that continues past the 8 core case
is due to cacheline pingponging when executing the rendevous. Both these sources
of latency drift could be corrected using replication quorums mapped to
dedicated NICs and more scalable rendezvous designs (such as with machine aware
broadcast trees~\cite{broadcast_tree}). However we deemed the complexity of such
optimizations unnecessary. The added latency for most cases is well under the
extra round trip delay, which would the minimum needed for a solution using
distributed transactions via two phase commit.

@@

We evaluate two different request sizes: 8 byte keys with 8 byte values and 8
bytes keys with 256 byte values. We run 8 physical logs on 8 cores and devote 32
cores to making requests to Rocksdb. Since we are interested in performance of
the log, which is only used for update requests, our workload is 100\% writes.

Before evaluating with Rocksdb we measure the baseline performance of
replicating the log with Cycloine for the given request sizes - Rocksdb performs
a no-op. We note that in addition to the key and value, we are also logging
RocksDB specific request data such as operation type and the request header.
Figure~\ref{fig:kv_baseline} shows the baseline performance for the chosen
request sizes. With the smaller request size, Cyclone can conservatively sustain
close to a million requests a second at a latency of just under 25us. With the
larger request size, Cyclone can sustain around 350K requests a second, again at
a latency of just under 25 us. Armed with these baseline numbers we now examine
how well Cyclone performs with Rocksdb.

@@

The performance of Rocksdb with Cyclone for the small update workload is shown
in Figure~\ref{fig:rocksdb}. We consider four different settings. The line
labeled Rocksdb is the key value store running with no logging whatsover - a
system crash would lead to data loss. The line labeled Rocksdb/WAL is for
Rocksdb running with its write ahead logging turned on. The large gap between
these two is the overhead of the existing Rocksdb WAL solution. The line labeled
Rocksdb/Cyclone 1 way is a two level Cyclone log but without any
replication. The line almost exactly tracks the performance of Rockdb. As
suggested by the baseline replication performance, Cyclone is able to provide a
write ahead log with no overhead to Rocksdb. The line labeled Rocksdb/Cyclone 3
way is with 3-way replication turned on. Other than a 20us delta due to the
extra network round trip, the line almost exactly tracks the Rocksdb no logging
case. Cyclone therefore provides high availability to Rocksdb at a fraction of
the cost of its existing single machine write ahead log. We also repeat the
experiment for the larger update size in Figure~\ref{fig:rocksdb_256}. The
conclusions are identical: Cyclone solves Rocksdb's write ahead logging
problem.

@@

Next, we consider the problem of supporting Rocksdb's write-batch operation that
atomically writes a set of key-value pairs into the KV store. In order to
perform the operation with Cyclone managing the log, the client must issue a
ganged operation across cores owning the keys in the write batch. This is in
contrast to baseline Rocksdb where the entire write batch can be sent to any
core. A key concern here was whether Cyclone would add any latency to the
operation due to the extra synchronization needed across replication quorums and
participating application cores. We examine the problem for the unloaded case
and small updates in Figure~\ref{fig:rocksdb_multi} for a single client with
increasing number of keys in the batch - till 32 keys that covers all
application cores. The line labeled Rocksdb is with no write-ahead logging. We
note an increasing latency for this baseline indicating Rocksdb takes longer
with larger key batches. For the other options - Rocksdb/wal has considerably
larger latency. Cyclone does an effective job of cutting down on this latency
even as it needs to pay a price for synchronizing multiple quorums and
application cores making it somewhat slower than running Rocksdb with no logging
for batched writes.

@@

Finally we showcase the benefit of using Cyclone beyond pure performance as
compared to the existing single machine Rocksdb write ahead log. Cyclone brings
multi-machine availability with the ability to automatically failover. We
demonstrate this in Figure ~\ref{fig:timeline} that shows the timeline of a run
where we kill the server process on the leader replica. Cyclone is configured to
with a 30ms failure detection timeout after which the client library tries
replicas in turn to locate the leader - in this case it fails over in about
60ms.

@@

Cyclone takes a software only approach to improving replication performance over
commodity network hardware. At the other end of the spectrum, Consensus in a
box~\cite{consensus_box} implements the zookeeper atomic broadcast protocol
together with a NIC on an FPGA. That work ignored durability focusing purely on
replication performance. Latency numbers are impressive from the FPGA that cuts
out the path from CPU caches to the NIC are as low as 3us for a round trip from
leader to follower replica compared to the 7us we observe with DPDK. Notably
however they reported 7us when using TCP/IP from the FPGA rather than their own
connection oriented network protocol. On the other hand peak replication
throughput appears to be 6M replications/sec on 66 Gbps of aggregate network
bandwith, comparable to the mark obtained by Cyclone on an aggregate 40 Gbps of
network bandwidth. We believe the general applicability afforded by our software
only packes switching solution makes it attractive in comparison to an FPGA
based solution even given the larger latency.

Networked Ordered Paxos~\cite{nopaxos} reprograms top of rack switches to order
client requests and multicast them to replicas - attacking replication latency
by removing the need to direct requests through a leader replica. It is hard to
compare performance due to the 1Gbps links used to leaf nodes in NoPaxos but we
note that avoiding changes to switching infrastructure was a specific constraint
in the design of Cyclone due to the difficulty we encountered in convincing
public cloud customers to allow such changes to be made.

{CRANE}~\cite{crane} is a system for replicating mutithreaded programs using
Paxos. It uses deterministic multi-threading to ensure replicas converge to the
same state as opposed to semantic equivalence in Cyclone using the commutativity
of the key value store interface. Although CRANE is more generally applicable in
that it makes no assumptions about the application API, deterministic
multithreading comes with high overheads for applications that store large
amounts of data - CRANE reports around a 2X slowdown for MySQL likely rendering
it unsuitable as a replacement for the WAL in key value stores.

A number of researchers have proposed using high performance networking hardware
for improving the performance of various distributed system stacks such as
distributed transactions~\cite{farm} and replication~\cite{dare, mojim}. We note
that the ideas in Cyclone are not specific to ethernet. The same ideas of packet
switching and horizontal scaling can be equally applied to RDMA. However, we
chose commodity ethernet due to the short term deployment plans that are a
priority design target.

The scalable commutativity rule~\cite{scalable_commutativity} generalized the
idea of commutativity in key value store interfaces that we have used for
horizontal scaling in Cyclone. Conveniently it provides a tool for checking when
API calls can commute in a history. It is a simple exercise to express a key
value store in COMMUTER and check that operations to different keys
commute. This provides a theoretical basis for horizontal scaling, something we
have omitted for brevity from this paper. The scalable commutativity rule also
suggests that scalable key value store implementations would exploit
commutativity something we have observed in the case of in-memory layers for key
value stores such as Flodb and indeed in the way we distributed keys across
physical logs in Cyclone.

Finally, we note the in progress transformation of durable memory in datacenters
with the advents of 3DXPoint. Researchers have speculated on the impact of cheap
and high volume directly attached 3DXPoint~\cite{tiering}. Memory vendors have
also begun shipping software libraries such as NVML~\cite{nvml} for building durable
in-memory data structures using undo logging. A durable key value store using
3DXPoint would be an idea target for Cyclone because each NVM log entry can be
discarded as soon as it is applied to the key value store, since the results are
persistent. This would allow Cyclone to operate at pure packet switching speeds,
a goal we hope to achieve in the future with a single level log for NVML
applications.

\section{Conclusion}

\newcommand\myurl[2]{\url{#1}}
\bibliographystyle{plain}
\bibliography{paper}

>>>>

% Paper number and no. of pages as author
\authorinfo{Paper \textbf{\#224}}{13 pages}

@@

In the absence of failures Figure~\ref{fig:ganged_ops} is straightforward. Once
replication is complete on all participating physical logs, the corresponding KV
store cores execute \\{\tt event\_replicated\_ganged\_op}. They use a bitmask for
the barrier setting the appropriate bit for themselves. Once all KV store cores
have arrived at the barrier the operation is executed. A distinguished leader
core (we use the minumum numbered participating KV store core) is
responsible for sending back the response to the client. Two types of failure
need to be considered. First, replication on the physical log corresponding to
one of the non-leader cores could fail. The
remaning cores would never progress past the barrier. To solve this the leader
core montors the remaning cores for their current view - which is the RAFT term
of the last executed log entry. If the view has moved past the view in the
packet, replication failed on that physical log. The leader then sets the bit
for the core that will no longer participate in the barrier to aid progress (a
technique we borrowed from lock-free data structures~\cite{lock-free}). It also
sets a failed flag to signal failure to the other participating cores. The
second failure case is that the leader core might not receive the operation if
it fails during replication from corresponding physical log. This case is
detected in a similar manner by the remaining cores by monitoring the published
view from the leader core in shared memory. We add that on a leadership change
due to timeout we enqueue and execute a nop on the physical log and all
associated KV store cores to ensure previous entries are committed - a corner
case in the RAFT protocol.

@@

\section{Evaluation}
\label{sec:evaluation}
We evaluate Cyclone on a 12 node x86 Xeon cluster connected via a 10 GigE
switch. Three of the machines are equipped with 1.6TB Intel DC P3600 SSDs and
4*10 GigE ports. The remaining nine machines do not have SSDs and have only one
10 GigE port, serving as clients for most of the experiments. We turn on jumbo
frame support in the 10 GigE switch to enable maximum use of batching in
Cyclone. As with other work~\cite{faast}, we use DRAM on the machines to proxy
for NVDIMMs where necessary - the persistent memory needed never exceeds 64 MB
regardless of the size of the key value store or second level log on flash. We
divide the evaluation into three parts. First, we evaluate Cyclone's performance
with a single level log as a pure software packet switch. Next, we evaluate
performance when adding a second level of log on flash. Finally, we evaluate
performance when integrated with Rocksdb~\cite{rocksdb} as an alternative to
Rocksdb's write ahead log. Unless otherwise mentioned, we use a 60 byte header
followed by an optional payload for experiments. We log both the header and
payload. In all cases the server echoes the received entry (header and payload)
back to the client. We fix the number of KV Store threads at 32 using at most 8
physcial logs with associated control plane cores for running RAFT instances.

Cyclone is built on top of DPDK to apply software packet switching techniques to
the log replication problem. We therefore begin by systematically evaluating
optimizations applied in Cyclone to replicate the NVM log in
Figure~\ref{fig:network_opts} - with no payload. The y-axis reports latency seen
at the client (which means two network round trips with replication). Using
TCP/IP to replicate a RAFT log tops out at around 30K entries/s.  Switching to
DPDK (the line marked +DPDK) improves the throughput by an order of magnitude to
around 500K entries/s. Using batching (the line marked +batching) improves the
performance further bringing us close to a million entries/s. Scaling
horizontally to 8 physical logs (+8 phy logs) improves performance to close to
2M entries/s. Finally using all 4 ethernet ports on the machine to replicate
entries improves performance considerably to 6M entries/s. In all, performance
improves by 200X over the TCP/IP single log baseline. Cyclone also considerably
improves the latency for replication, from close to 100us with TCP/IP to around
30us at peak throughput.

@@

There are two factors that can have significant impact on Cyclone's
performance. First, the number of replicas dictates the outgoing message rate
from the leader replica and therefore increasing the replication factor can
decrease Cyclone's performance. Figure~\ref{fig:replicas} shows the impact of
varying replica count. Using only a single replica cuts out a network round trip
and shows the best unloaded latency (10 us) and peak throughput (near 10M
entries/s). Adding replicas decreases the peak throughput down to around 2M
entries/s with 5 replicas. We note that a number of previous pieces of
work~\cite{faast, farm} use three replicas and therefore we focus on three
replicas for the replicated cases we consider below. The second factor that
dictates Cyclone's performance is the size of the log entry being
replicated. Figure~\ref{fig:payload} shows the effect of increasing the payload
size from zero to 512 bytes. Peak throughput drops from 6M entries/s to
approximately 2M entries/s. At this replication rate, the leader replica needs
to transmit data at approximately 30 Gbit/s. Coupled with the cost of network
headers all four 10 GigE ports are now saturated and therefore Cyclone hits the
network line rate bottleneck at this point.

@@

We now turn our attention from the network component of Cyclone to the storage
one by adding the second level flashlog. We evaluate the impact of adding
block storage to our hitherto pure packet switching scenario in
Figure~\ref{fig:flashlog}. Batching entries from the top level NVDIMM log to the
second level flashlog is clearly beneficial as adding flash storage at the
second level has almost no impact on peak performance in Cyclone for small
entries. The situation however changes for larger
entries. Figure~\ref{fig:flashlog_512} shows that using a 512 byte payload has a
significant impact on peak throughput - it drops to approximately 350K
ops/sec. This corresponds to around 50K 4KB IOPS to the SSD to write out the
flashlog pages. The peak for the drive is 160K IOPS using a queue depth
(concurrency) of 128. With 32 application threads we expect a lower peak
throughput around 40K IOPS explaining our bottleneck at 50K IOPS.  It is
possible to tune our observed performance further by aligning the flush boundary
to increase the number of outstanding requests - we do not do so in this paper,
keeping Cyclone agnostic to the exact size of entry being replicated. A final
point about Figure~\ref{fig:flashlog_512} is that once we are past the storage
bottleneck the latency spike is dramatic and large enough to trigger Cyclone's
failure detector and repeated retries from the clients. There are - therefore -
no points on the ``knee'' of the curve as in the pure packet switched one level
log case.

@@

\begin{figure}
\includegraphics[scale=0.6]{results2/flashlog_512.pdf}
\caption{Second level log with 512 bytes payload}
\label{fig:flashlog_512}
\end{figure}

The final dimension we evaluate is of ganged operations. The primary purpose of
ganged operations is to avoid the need for distributed transactions to
manipulate what is a single shared memory image and therefore we were most
concerned about unloaded latency given the complexity of synchronizing different
replication quorums as well executing our rendezvous protocol on multiple
cores. We therefore setup an experiment where a single client - reflecting the
unloaded case - made ganged requests to the replicas. We varied the number of
cores participating in the request from one to the full complement of 32
application cores. Figure~\ref{fig:ganged} shows the results both using a single
level log as well as a two level log. The primary takeaway is that unloaded
latency increases slowly as we increase the number of active cores - to around
40 us from the baseline of 20 us. There are two causes of this. First, there is
a quicker rate of increase from 1 to 8 active cores, the reason being that this
corresponds to an increase in the number of physical logs that must be
synchronized on. There is some amount of synchronization involved in the
userspace DPDK driver for access to common resources for accessing the NIC and
this affects the ability to simultaneously issues replication messages on all
quorums. The smaller contribution to increasing latency that continues past the
8 core case is due to cacheline pingponging when executing the rendevous of
Figure~\ref{fig:ganged_ops}. Both these sources of latency drift could be
corrected using replication quorums mapped to dedicated NICs and more scalable
rendezvous designs (such as with machine aware broadcast
trees~\cite{broadcast_tree}). However we deemed the complexity of such
optimizations unnecessary. The added latency for most cases is well under the
extra round trip delay (including replication for all phases of a 2PC), which
would the minimum needed for a solution using distributed transactions via two
phase commit.

@@

We evaluate two different request sizes: 8 byte keys with 8 byte values and 8
byte keys with 256 byte values. We run 8 physical logs on 8 cores and devote 32
cores to Rocksdb. Since we are interested in performance of the log, which is
only used for update requests, our workload is 100\% writes.

Before evaluating with Rocksdb, we measure the baseline performance of
replicating the log with Cyclone for the given request sizes - Rocksdb performs
a no-op. We note that in addition to the key and value, we are also logging
RocksDB specific request data such as operation type and the request header.
Figure~\ref{fig:kv_baseline} shows the baseline performance for the chosen
request sizes. With the smaller request size, Cyclone can conservatively sustain
close to a million requests a second at a latency of just under 25us. With the
larger request size, Cyclone can sustain around 350K requests a second, again at
a latency of just under 25 us. Armed with these baseline numbers we now examine
how well Cyclone performs with Rocksdb.

@@

The performance of Rocksdb with Cyclone for the small update workload is shown
in Figure~\ref{fig:rocksdb} - essentially presenting the solution to the problem
description graph in Figure~\ref{fig:problem}.  We consider four different
setups. The line labeled 'rocksdb' is the key value store running with no
logging whatsover - a system crash would lead to data loss. The line labeled
'rocksdb/wal' is for Rocksdb running with its write ahead logging turned on. The
large gap between these two is the overhead of the existing Rocksdb WAL
solution. The line labeled 'rocksdb/Cyclone 1 way' is a two level Cyclone log
but without any replication. The line almost exactly tracks the performance of
Rockdb. As suggested by the baseline replication performance, Cyclone is able to
provide a write ahead log with no overhead to Rocksdb. The line labeled
'rocksdb/Cyclone 3 way' is with 3-way replication turned on. Other than a 20us
delta due to the extra network round trip, the line almost exactly tracks
Rocksdb performance with no logging. Cyclone therefore provides high
availability to Rocksdb at a fraction of the cost of its existing single machine
write ahead log. We also repeat the experiment for the larger update size in
Figure~\ref{fig:rocksdb_256}. The conclusions are identical: Cyclone solves
Rocksdb's write ahead logging problem.

@@

Next, we consider the problem of supporting Rocksdb's write-batch operation that
atomically writes a set of key-value pairs into the KV store. In order to
perform the operation with Cyclone managing the log, the client must issue a
ganged operation across cores owning the keys in the write batch. This is in
contrast to baseline Rocksdb where the entire write batch can be sent to any
core. A key concern here was whether Cyclone would add any latency to the
operation due to the extra synchronization needed across replication quorums and
participating application cores. We examine the problem for the unloaded case
and small updates in Figure~\ref{fig:rocksdb_multi} for a single client with
increasing number of keys in the batch - till 32 keys that covers all
application cores. The line labeled Rocksdb is with no write ahead logging. We
note an increasing latency for this baseline indicating Rocksdb takes longer
with larger key batches. The existing option of Rocksdb/wal has considerably
larger latency. Cyclone does an effective job of cutting down on this latency
even as it needs to pay a price for synchronizing multiple quorums and
application cores making it somewhat slower than running Rocksdb with no logging
for batched writes. A distributed 2PC transaction still remains more expensive
just in terms of network delay than using Cyclone.

@@

Finally, we showcase the benefit of using Cyclone beyond pure performance as
compared to the existing single machine Rocksdb write ahead log. Cyclone brings
multi-machine availability with the ability to automatically failover. We
demonstrate this in Figure ~\ref{fig:timeline} that shows the timeline of a run
where we kill the server process on the leader replica. Cyclone is configured to
with a 30ms failure detection timeout after which the client library tries
replicas in turn to locate the leader - in this case it fails over in about
60ms.

@@

Cyclone takes a software only approach to improving replication performance
using commodity network hardware. At the other end of the spectrum, Consensus in
a box~\cite{consensus_box} implements the Zookeeper atomic broadcast protocol
together with a NIC on an FPGA. That work ignored durability focusing purely on
replication performance. Latency numbers are impressive from the FPGA that cuts
out the path from CPU caches to the NIC and are reported to be as low as 3us for
a round trip from leader to follower replica compared to the 7us we observe with
DPDK. Notably however they reported 7us when using TCP/IP from the FPGA rather
than their own connection oriented network protocol. On the other hand peak
replication throughput appears to be 6M replications/sec on 66 Gbps of aggregate
network bandwith, comparable to the mark obtained by Cyclone on an aggregate 40
Gbps of network bandwidth. We believe the general applicability afforded by our
software only packet switching solution makes it a compelling alternative to an
FPGA based solution even given the larger latency.

Network Ordered Paxos~\cite{nopaxos} reprograms top of rack switches to order
client requests and multicast them to replicas - attacking replication latency
by removing the need to direct requests through a leader replica. It is hard to
compare performance due to the 1Gbps links used to leaf nodes in NoPaxos but we
note that avoiding changes to switching infrastructure was a specific constraint
in the design of Cyclone due to the difficulty we encountered in convincing
public cloud customers to allow such changes to be made.

{CRANE}~\cite{crane} is a system for replicating mutithreaded programs using
Paxos. It uses deterministic multi-threading to ensure replicas converge to the
same state as opposed to semantic equivalence in Cyclone using the commutativity
of the key value store interface. Although CRANE is more generally applicable in
that it makes no assumptions about the application API, deterministic
multithreading comes with high overheads for applications that store large
amounts of data - CRANE reports around a 2X slowdown for MySQL likely rendering
it unsuitable as a replacement for the write ahead log in key value stores.

A number of researchers have proposed using high performance networking hardware
for improving the performance of various distributed system primitives such as
distributed transactions~\cite{farm} and replication~\cite{dare, mojim}. We note
that the ideas in Cyclone are not specific to ethernet. The same ideas of packet
switching and horizontal scaling can be equally applied to RDMA. However, we
chose commodity ethernet due to short term deployment plans that are a
priority design target.

The scalable commutativity rule~\cite{scalable_commutativity} generalizes the
idea of commutativity in key value store interfaces that we have used for
horizontal scaling in Cyclone. Conveniently, it provides a tool for checking when
API calls can commute in a history. It is a simple exercise to express a key
value store API in COMMUTER and check that operations to different keys
commute. This provides a theoretical basis for horizontal scaling, something we
have omitted for brevity from this paper. The scalable commutativity rule also
suggests that scalable key value store implementations would necessarily have
commutative interfaces, something we confirm and exploit for key
value stores.

There is an in progress transformation of durable memory in datacenters
with the advents of 3DXPoint. Researchers have speculated on the impact of cheap
and high volume directly attached 3DXPoint~\cite{tiering}. Memory vendors have
also begun shipping software libraries such as NVML~\cite{nvml} for building durable
in-memory data structures using undo logging. A durable key value store using
3DXPoint would be an idea target for Cyclone because each NVM log entry can be
discarded as soon as it is applied to the persistent key value store removing
the need for the flashlog. This would allow Cyclone to operate at pure packet
switching speeds, a goal we hope to achieve in the future with a single level
log for NVML data store applications.

Cyclone provides a client library that automatically times out and retries
requests to deal with failure. Protection against repeated execution of the same
request if currently left to the key value store and client, since we felt that
exactly once semantics is optional when updates such as put operations are
idempotent. We have experimented with exactly once semantics as a basis for
distributed transactions across shards running Cyclone integrated key value
stores, along the lines of similar work~\cite{raft_lin}, using a fixed client
space and maintaining information about last seen sequence numbers from clients
in NVM. Distributed transactions on Cyclone shards however are outside the scope
of this paper.

\section{Conclusion}
Persistent key values stores struggle today to provide high availability using a
write ahead log without a significant impact to performance. Cyclone solves this
problem by leveraging a small amount of directly attached non-volatile memory
that transforms the replication problem into a software packet switching
problem. Cyclone and its Rocksdb bindings are slated for release to the open
source community and we hope both the code and the ideas in the system itself
will help key value store builders and researchers deal with their logging
problem.
\newcommand\myurl[2]{\url{#1}}
\bibliographystyle{plain}
\bibliography{paper}
